---
title: "Data Visualization"
author: "Jefferson Omokoh"
date: "5/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- ggplot2 implements the grammar of graphics

- ggplot2 is one of the core members of the tidyverse package

Question: Do cars with big engines use more fuel than cars with small engines?

*Dataframe: is a rectangular collection of variables (in the columns) and observation (in the rows).* 

```{r}
mpg
?mpg
```
```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point()
```
The above graph shows a positive relationship between engine size(displ) and fuel efficiency (hwy), cars with big engines use more fuel. 

- geom_point() adds a layer of points to my plot
- ggplot2 comes with many geom functions that each add a different type of layer to a plot

*A graphing template*
ggplot(data = <DATA>) +
  GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
  
```{r}
mpg %>% glimpse()

ggplot(mpg, aes(x = class, y = drv)) +
  geom_point()
```
  
- The above graph is not useful because it does not give information about a relationship that exist.

*Aesthetic mappings*
- I can add a third variable, like class, to a two dimensional scatterplot by mapping it to an aesthetic.
- Aesthetic: is a visual property of the objects in my plot. Aesthetics include things like the shape, the size, or the color of my points.
- **I can convey information about my data by mapping the aesthetics in my plot to the variables in my dataset**

```{r}
ggplot(mpg, aes(x = displ, y = hwy, color =  class)) +
  geom_point()
```
To map an aesthetic to a variable, associate the name of the aesthetic to the name of the variable inside aes(). ggplot2 will automatically assign a unique level of the aesthetic (here a unique color) to each unique value of the variable, a process known as *scaling*. ggplot2 will also add a legend that explains which levels correspond to which values.

- We can also map class to the size aesthetic, see below

```{r}
ggplot(mpg, aes(x = displ, y = hwy, size = class)) + 
  geom_point()
```
- We can also use the alpha aesthetic, which controls the transparency of the points, or to the shape aesthetic, which controls the shape of the points

```{r}
ggplot(mpg, aes(x = displ, y = hwy, alpha = class)) +
  geom_point()

ggplot(mpg, aes(x = displ, y = hwy, shape = class)) +
  geom_point()
```
- Alpha assign different transparency to the different classes, while shape assigns different shapes to the different classes. Note: the shape aesthetic can only take 6 shapes at a time, and by default the additional groups will go unplotted.


- Setting the aesthetic property of my geom manually.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(color = 'blue')
```
- To set an aesthetic manually, set the aesthetic by name as an argument of geom function, i.e it goes outside of aes()


*Facets*
Facets: subplots that each display one subset of data
- To facet your plot by a single variable, use facet_wrap()
- The first argument in a facet_wrap should be a formula, which you create with ~ followed by a variable name.
- The variable that I pass to facet_wrap() should be discrete

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() +
  facet_wrap(~ class, nrow = 2)
```
- To facet my plot on the combination of two variables, add facet_grid() to the plot call.

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() +
  facet_grid(drv ~ cyl)
```
- If I would rather not facet in the rows or columns dimension, I can use a . instead of a variable name, e.g. + facet_grid(. ~ cyl). 


*Geometric objects*
- We can use different visuals in ggplot by using different geoms.
*geom:* is a geometrical object that a plot uses to represent data. 


```{r}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_smooth()
```

****
```{r}
ggplot(mpg, aes(x = displ, y = hwy,linetype = drv)) +
  geom_smooth()
```

- in the above plot geom_smooth() separates the cars into three lines based on their drv value, which describes a car's drivetrain.


```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
              
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
    
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )
```






